---
title: "sdcLog"
subtitle: "Toolkit for Output Control in Research Data Centres"
author: "Matthias Gomolka"
institute: "Deutsche Bundesbank, Research Data and Service Centre (RDSC)"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: yes
      highlightSpans: yes
      countIncrementalSlides: no
      slideNumberFormat: 'Slides available on <a href="https://tinyurl.com/sdcLog-presentation">https://tinyurl.com/sdcLog-presentation</>'
lang: de-de
editor_options: 
  chunk_output_type: console
---

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
library(sdcLog)
library(DiagrammeR)
style_mono_accent(
  base_color = "#1d3557",
  header_font_google = google_font("Fira Sans"),
  text_font_google   = google_font("Fira Sans", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
options(datatable.print.keys = FALSE)
options(datatable.print.class = FALSE)

knitr::opts_chunk$set(
  # prompt = TRUE,
  # comment = "#>",
  collapse = TRUE,
  background = "#FAFAFA"
)
```

# Who am I? 
### And why do I talk about sdcLog?

I work in the Bundesbank's Research Data and Service Centre.

--

#### What I do: 

--

- Production of research data sets on securities transactions
- Data Production Pipelines
- **R tools, which make the RDSC life easier**

--

#### Disclaimer:

- No expert in Output Control. I just implemented functionality which we already
  have for Stata in R.

---

# Motivation

#### Problem

- Researchers need to show that their output complies to our rules.

--

- That get's complicated quickly.

--

- It would be very time-consuming for the RDSC if we would have to check *how* a
  researcher proved that her output complies to RDSC rules.

--

#### Solution

- RDSC provides tools which help researchers to show that their output complies 
  to the rules: **sdcLog**

---

# Theory

Two simple rules:

--

1. Each result must be based on at least 5 distinct entities (distinct ID's).

--

2. The two largest entities must not account for more than 85% of a result
   (n,k-dominance).


---

# Example

A researcher wants to publish the mean of a variable grouped by `sector`. To do 
so, she has to use `sdc_descriptives()` to show that the output complies to RDSC
rules.

--

.pull-left[
```{r sdc_descriptives_1, echo=3}
data("sdc_descriptives_DT")
DT <- sdc_descriptives_DT[, id_na := NULL]
head(DT)
```
]

--

.pull-right[
```{r sdc_descriptives_3}
# result
DT[, .(mean = mean(val_1, na.rm = TRUE)),
   by = "sector"]
```
]

--

```{r sdc_descriptives_6}
# Proof, that the result complies to rules
sdc_descriptives(DT, id_var = "id", val_var = "val_1", by = "sector")
```

---

# Another example

This time, researches want to calculate the result grouped by `sector` and
`year`.

--

```{r sdc_descriptives_7}
sdc_descriptives(DT, id_var = "id", val_var = "val_1", by = c("sector", "year"))
```

---

# Minimum and maximum values

Now, researchers want to publish minimum and maximum values as well. 

--

#### Problem

Minimum and maximum value are confidential micro data.

--

#### Solution

"Minimum" and "maximum" value as mean of `n` smallest / largest values using 
`sdc_min_max()`:

--

```{r sdc_min_max_1}
sdc_min_max(DT, id_var = "id", val_var = "val_1")
```

---

# Output control for models

Researchers also want to publish results from a linear regression.

--

```{r sdc_model_1}
options(sdc.n_ids = 3)

# Estimate model
mod <- lm(val_1 ~ sector + year + val_2, data = DT)

# Check if model complies to rules
sdc_model(DT, model = mod, id_var = "id")
```

---

# Why is it called sdcLog?

<center>
```{r diagrammer_1, echo=FALSE, fig.align='center', out.width='60%'}
grViz("
digraph boxes_and_circles {

  # a 'graph' statement
  graph [overlap = true, fontsize = 30]
  
  # several 'node' statements
  node [shape = box, fontname = 'Fira Sans', width = 5.5]
  script [label = 'analysis.R\ncontains sdc_descriptives() / sdc_min_max() / sdc_model()'];
  
  node [width = 1] 
  log [label = 'Log file']
  
  node [shape = box, width = 3]
  checks [label = 'Checked by RDSC']

  node [shape = oval, width = 1] 
  sdc_log [label = '@@1']

  # several 'edge' statements
  script -> sdc_log
  sdc_log -> log
  log -> checks
}

[1]: 'sdc_log(analysis.R)'
")
```
</center>

---

# Installation und contact information

#### CRAN

```{r install, eval=FALSE}
install.packages("sdcLog")
```

#### GitHub

https://github.com/matthiasgomolka/sdcLog/issues

#### E-mail

[matthias.gomolka@bundesbank.de](mailto:matthias.gomolka@bundesbank.de)

#### Phone

069 9556-4991
